---
title: "Covid91 vaccine study Final2021"
author: "Carly Syms"
date: "May 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods 

The first step is isolating the sub-population 

```{r}
maledata<- filter(FinalData,sex=="guy")
maledata$sex<- as.factor(maledata$sex)
maledata$infected=factor(maledata$infected)
maledata$infected<-relevel(maledata$infected,"OK")
levels(maledata$infected)
```
```{r}
femaledata<- filter(FinalData,sex=="gal")
femaledata$sex<- as.factor(femaledata$sex)
femaledata$infected=factor(femaledata$infected)
femaledata$infected<-relevel(femaledata$infected,"OK")
levels(femaledata$infected)
```
```{r}
druggiedata<- filter(FinalData,DrugUser=="yes")
druggiedata$DrugUser<- as.factor(druggiedata$DrugUser)
druggiedata$infected=factor(druggiedata$infected)
druggiedata$infected<-relevel(druggiedata$infected,"OK")
levels(druggiedata$infected)
```

Then we will look at the Barplots, Numerical Summaries, Inferential Results, and Fisher Exact Tests of each sub-population. 

# Males 

## Barplot

```{r}

Males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart shows that out of the approximately 340 male who were infected with Covid, 57% were given the drug, and 43% were given the placebo.

## Numerical Summery

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 males were infected with Covid out of the approximately 10,400 males in the study. The table shows that about 3.72% of the males who received the drug got Covid, and 2.55% of the males who received the placebo got Covid. It also shows that the males who got the drug were about 50% more likely to contract Covid. This suggests that the drug is ineffective for males. 

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared table shows that the p-value of 0.0006181 is so small that it is highly unlikely that the results are by chance.  
$H_O$ the null hypothesis given taking the vaccine there is no difference of whether or not a male can contracts Covid.  
$H_A$ the alternative hypothesis is that the vaccine does make a difference in whether or not a male can contract Covid. Therefor we reject the Null hypothesis.   
The odds ratio in the fisher exact test is 1.47:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 59.5% effectiveness. 

# Females

## Barplot

```{r}

Females <- femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart shows that out of the approximately 610 female who were infected with Covid, 33% were given the drug, and 66% were given the placebo.

## Numerical Summery

```{r}
table2 <- xtabs(~treatment + infected, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

The table shows that about 610 females were infected with Covid out of the approximately 9,200 females in the study. The table shows that about 4.44% of the females who received the drug got Covid, and 8.67% of the females who received the placebo got Covid. It also shows that the females who got the placebo were about 50% more likely to contract Covid. This suggests that the drug is effective for females. 

## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The Chi Squared table shows that the p-value of 3.626e-16 is so small that it is highly unlikely that the results are by chance.    
The odds ratio in the fisher exact test is 0.49:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 67% effective for women.

# Drug Users

## Barplot

```{r}

Druggies <- druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This bar chart shows that out of the approximately 117 drug users who were infected with Covid, 47% were given the drug, and 53% were given the placebo.

## Numerical Summery

```{r}
table3 <- xtabs(~treatment + infected, data=druggiedata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

The table shows that about 117 drug users who were infected with Covid out of the approximately 2,200 drug users in the study. The table shows that about 3.63% of the drug users who received the drug got Covid, and 6.46% of the drug users who received the placebo got Covid. This suggests that the drug is effective for drug users. 

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The Chi Squared table shows that the p-value of 0.003028 is so small that it is highly unlikely that the results are by chance.    
The odds ratio in the fisher exact test is 1.83:1, when taking the odds ratio and computing it into a percent of effectiveness by using the formula 0/(1+0)*100 you get 60% effectiveness. 

# LGBTQ 



# Discussion and Conclusion  
